\textbf{Selective DNA and Protein Isolation from Marine Macrophyte Surfaces}

### 2019-12-02 submission/manuscript.Rmd
Rendering optimization

## Figures

```{r out.width="100%", fig.align="center", fig.cap="Bacteril Community\\label{community}", fig.pos="h"}
include_graphics("../results/figures/community_barplot.jpg")
```

>>>>

## Figures
```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="h"}
include_graphics("../results/figures/community_bar_plot.jpg")
```

### 2019-12-02 submission/manuscript.tex
Customizing .Rprofile
Rendering optimization

@@

\hypertarget{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}{%
\subsubsection{DNA Isolation, Illumina 16S rRNA Sequencing and Sequences
Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}}

@@

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing, at IMGM the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a postive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. The sequences obtained in this study have been submitted to the
European Nucleotide Archive (ENA) under accession numbers ?????.

\begin{verbatim}
## Parsed with column specification:
## cols(
##   .default = col_double(),
##   query = col_character(),
##   reference = col_character()
## )
\end{verbatim}

\begin{verbatim}
## See spec(...) for full column specifications.
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   taxlevel = col_double(),
##   rankID = col_character(),
##   taxon = col_character(),
##   daughterlevels = col_double(),
##   total = col_double(),
##   `40` = col_double(),
##   `41` = col_double(),
##   `42` = col_double(),
##   `43` = col_double(),
##   `61` = col_double(),
##   `62` = col_double(),
##   `63` = col_double(),
##   ATCC_1 = col_double(),
##   NC_1 = col_double()
## )
\end{verbatim}

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommandations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplsts by
mothur were exported, aligned using SILVA incremental aligner (SINA,
version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
(Ludwig \emph{et al.}, 2004) for further phylogentic analysis using the
same database. Reference sequences close to imported ones were selected
and used to calculate a phylogentic tree using the Maximum Likelihood
algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
2006). Imported partial chloroplast sequences were added to the tree
using the maximum parsimony criteria and not allowing changes to tree
topology. Pipeline data processing and visualizaion was done using R
(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
al.}, 2019). The detailed analysis procedure including the R Markdown
file for this paper are available as a GitHub repository (?????). Based
on the ATCC MSA-1002 mock community included in the analysis a
sequencing error rate of 0.009 \si{\percent} was detemined that is in
line with previously reported values for next-generation sequencing data
(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
the negative control processed together with the samples yielded only
44,050 sequences.

\hypertarget{protein-isolation-and-metaproteomics}{%
\subsubsection{Protein Isolation and
Metaproteomics}\label{protein-isolation-and-metaproteomics}}

Proteins were isolated according to the protocol for isolation from soil
described in Chourey \emph{et al.} (2010) and modified by Hultman
\emph{et al.} (2015). These protocols were further modified and adapted
for protein isolation from microbial communities form macrophyte
surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
thalli. The mixture was incubated in boiling water for 5 \si{\minute},
vortexed for 10 \si{\minute} and incubated again in boiling water for 5
\si{\minute}. After a brief vortex the lysate was transferred to a clean
tube separating the host leaves or thalli from the mixture containing
lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
\si{\milli\Molar}) was added and proteins were precipitated with chilled
100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
The obtained protein pellet was washed three times with chilled acetone.
During the first washing step the pellet was transferred to a clean 1.5
\si{\ml} tube. After each washing step the samples were centrifuged at
20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
were stored at \num{-80} °C until further analysis.

@@

\hypertarget{results-and-discussion}{%
\subsection{Results and Discussion}\label{results-and-discussion}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\hypertarget{materials-and-methods-1}{%
\subsection{Materials and Methods}\label{materials-and-methods-1}}

@@

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

@@

@@

>>>>

\usepackage{caption}
\captionsetup{justification=justified}

@@

\hypertarget{dna-isolation}{%
\subsubsection{DNA Isolation}\label{dna-isolation}}

@@

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
Sequencing}\label{illumina-16s-rrna-sequencing}}

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing, at IMGM the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. The sequences obtained in this study have been submitted to the
European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
ADDED LATER!}.

\hypertarget{sequence-analysis}{%
\subsubsection{Sequence Analysis}\label{sequence-analysis}}

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
mothur were exported, aligned using SILVA incremental aligner (SINA,
version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
same database. Reference sequences close to imported ones were selected
and used to calculate a phylogenetic tree using the Maximum Likelihood
algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
2006). Imported partial chloroplast sequences were added to the tree
using the maximum parsimony criteria and not allowing changes to tree
topology. Pipeline data processing and visualization was done using R
(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
al.}, 2019). The detailed analysis procedure including the R Markdown
file for this paper are available as a GitHub repository (\textbf{TO BE
ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
the analysis a sequencing error rate of 0.009 \si{\percent} was
determined, which is in line with previously reported values for
next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
\emph{et al.}, 2016). In addition, the negative control processed
together with the samples yielded only 2 sequences after sequence
quality curation.

\hypertarget{protein-isolation}{%
\subsubsection{Protein Isolation}\label{protein-isolation}}

Proteins were isolated according to the protocol for isolation from soil
described in Chourey \emph{et al.} (2010) and modified by Hultman
\emph{et al.} (2015). These protocols were further modified and adapted
for protein isolation from microbial communities form macrophyte
surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
thalli. The mixture was incubated in boiling water for 5 \si{\minute},
vortexed for 10 \si{\minute} and incubated again in boiling water for 5
\si{\minute}. After a brief vortex the lysate was transferred to a clean
tube separating the host leaves or thalli from the mixture containing
lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
\si{\milli\Molar}) was added and proteins were precipitated with chilled
100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
The obtained protein pellet was washed three times with chilled acetone.
During the first washing step the pellet was transferred to a clean 1.5
\si{\ml} tube. After each washing step samples were centrifuged at
20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
were stored at \num{-80} °C until further analysis.

\hypertarget{metaproteomics}{%
\subsubsection{Metaproteomics}\label{metaproteomics}}

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions (Wiśniewski \emph{et al.},
2009) with small modifications. Before loading the solution to the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing digested proteins was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid, freezed at
\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
to the Vienna Metabolomics Center (University of Vienna) for
metaproteomic analysis. \textbf{TO BE ADDED-VIENNA PART}.

@@

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{} 3 \si{\percent}). In addition,
nontreated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For long therm storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

@@

\hypertarget{references}{%
\subsection{References}\label{references}}

@@

\leavevmode\hypertarget{ref-Wisniewski2009}{}%
Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
Universal sample preparation method for proteome analysis. \emph{Nature
Methods} \textbf{6}: 359--362.

@@

\hypertarget{figure-captions}{%
\subsection{Figure Captions}\label{figure-captions}}

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[h]

{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-2}
\end{figure}
